\documentclass[11pt]{book}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[margin=1.3in]{geometry}

%For Baskerville if wanted
\usepackage{kpfonts,baskervald}
\usepackage[multiple]{footmisc}
\usepackage{cancel}
\usepackage{float}
\usepackage{comment}
\usepackage{mathtools}
\usepackage{etoolbox}

%Bib style
\bibliographystyle{jhep}
\usepackage{cite}

%For better equations and such.
\usepackage{amsmath}
\usepackage{stmaryrd}
\usepackage{physics}

%For tables.
\usepackage[table,xcdraw]{xcolor}
\usepackage{enumitem}
\usepackage{slashed}
\usepackage{multirow}

%For better quivers
\usepackage[all]{xy}

%For tikz and plots
\usepackage{adjustbox}
\usepackage{tikz}
\usetikzlibrary{positioning,
				cd,
				decorations.markings, 
				arrows.meta,
				tqft,
				calc}
\usepackage{pgfplots}
\pgfdeclarelayer{background layer}
\pgfdeclarelayer{foreground layer}
\pgfsetlayers{background layer,main,foreground layer}
\usepackage[labelfont=bf]{caption}
\usepackage{blindtext}
\pgfplotsset{compat=1.17} 

\usepackage{tcolorbox}
\tcbuselibrary{theorems}
\newtcbtheorem[number within=section]{defn}{Definition}%
{colback=green!5,colframe=green!35!black,fonttitle=\bfseries}{df}
\newtcbtheorem[number within=section]{thm}{Theorem}%
{colback=green!5,colframe=red!35!black,fonttitle=\bfseries}{th}
\newtcbtheorem[number within=section]{Jargon}{Jargon}%
{colback=green!5,colframe=blue!35!black,fonttitle=\bfseries}{jr}
%For numbered lemmas, theorems etc.
\usepackage{amsthm}
\theoremstyle{definition}
\newtheorem{example}{Example}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{conjecture}{Conjecture}[section]
\newtheorem{rem}{Remark}
\newtheorem{claim}{Claim}

%For unnumbered remarks.
\newtheorem*{remark}{Remark}

%For better numbering of equations.
\numberwithin{equation}{section}


%For clickable table of contents
\usepackage{hyperref}
\definecolor{airforceblue}{rgb}{0.36, 0.54, 0.66}
\hypersetup{
    colorlinks,
    citecolor=purple,
    filecolor=black,
    linkcolor=airforceblue,
    urlcolor=blue
}
\usepackage[nameinlink]{cleveref}

%%%%%%%

\newcommand{\flux}[2][]{[\rho_{#2}^{#1}]^{y_S}_{y_N}}
\NewDocumentCommand{\fluxx}{ O{} O{} m m}{[\rho_{#3}^{#1}\rho_{#4}^{#2}]^{y_S}_{y_N}}

\usepackage[bbgreekl]{mathbbol}

\usepackage{bbold}
\DeclareSymbolFontAlphabet{\mathbb}{AMSb}
\DeclareSymbolFontAlphabet{\mathbbl}{bbold}

\newcommand{\Spindle}{\mathbbl{\Sigma}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%Handy macros%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\renewcommand*\rm[1]{\mathrm{#1}}

%For cohomology group macros
\DeclareMathOperator{\HH}{H}
\DeclareMathOperator{\cHH}{\check{H}}
\DeclareMathOperator{\Hom}{Hom}


%For correct infinitesimals.
\newcommand*\diff{\mathop{}\!\mathrm{d}}
\newcommand*\Diff{\mathop{}\!\mathrm{D}}

%For correct imaginary unit
\newcommand*\I{\mathrm{i}}
\newcommand*\ee{\mathrm{e}}

%Letters
\usepackage{dowith}
\newcommand{\UpperCase}[1]{
  \expandafter\newcommand\csname bb#1\endcsname{\mathbb{#1}}
  \expandafter\newcommand\csname c#1\endcsname{\mathcal{#1}}
  \expandafter\newcommand\csname Fr#1\endcsname{\mathfrak{#1}}   }
\DoWith\UpperCase ABCDEFGHIJKLMNOPQRSTUVWXYZ\StopDoing

\newcommand{\LowerCase}[1]{
  \expandafter\newcommand\csname Fr#1\endcsname{\mathfrak{#1}}   }
\DoWith\LowerCase abcdefghijklmnopqrstuvwxyz\StopDoing

%Common script letters
\newcommand*\scrt{\mathscr{T}}
\newcommand{\ab}[1]{#1^{\mathrm{ab}}}
\newcommand{\GL}[1]{\mathrm{GL}(#1)}
\newcommand{\PGL}[1]{\mathrm{PGL}(#1)}
\newcommand*\U{\mathrm{U}}
\newcommand*\SU{\mathrm{SU}}
\newcommand*\PSU{\mathrm{PSU}}
\newcommand*\SO{\mathrm{SO}}
\newcommand*\OO{\mathrm{O}}
\newcommand*\SL{\mathrm{SL}}
\newcommand*\USp{\mathrm{USp}}
\newcommand*\Alt{\mathrm{Alt}}
\newcommand*\Sym{\mathrm{Sym}}
\newcommand*\bbCt{\mathbb{C}^\times}
\newcommand*\QFT{\mathbf{QFT}}
\newcommand{\ddv}{\mathrm{d}_V}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{\vspace{1cm}\bf{Various Notes}\vspace{0.2cm}}
\date{\vspace{0.2cm} \today}
\author{Davide Morgante}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\allowdisplaybreaks
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\setcounter{page}{1}
{\hypersetup{linkcolor=orange}
\tableofcontents
}
\mainmatter
\include{files/geom_gauge}
\include{files/spindles}
\include{files/gen_symm}
\include{files/cs_theory}
\include{files/power_sym}
%\include{files/rep_catsym}
\include{files/SW_theory}
\include{files/class_S}
\include{files/2d4d_corr}
\include{files/Lens_space}

\chapter{Various concepts}
\section{Topological twist}
SCFTs in low dimensions provide an interesting test of AdS/CFT since they describe the field theory dual of supergravity solutions in AdS backgrounds. In low dimensions, there are not many constraints from supersymmetry, therefore a full classification of SCFTs is not possible. Nevertheless it is still an open problem to find new ways to explore the space of such theories and find possible relations between the known ones.

SCFTs in $2d$ can be obtained by reducing $4d$ SCFTs on compact two-manifold. So we put the theory on a curved background $\bbR^{1,1}\times\Sigma_g$, where $\Sigma_g$ is a compact, closed Riemann surface of genus $g$. The spin connection $\omega$ has in general a non-trivial holonomy around $\Sigma$ which means that the fields that are coupled to it could transform non-trivially under susy transformations. For example, the susy variation of the gravitino field, i.e. the (odd) one-form connection associated to the supercharges, contains the covariant derivative of the susy parameters. These parameters are sections of the spinor bundle and as such they couple to $\omega$. If we want to preserve susy on such parameters, than they must be covariantly constant so that the gravitino variation vanishes. Spinors with this property usually only exist for manifolds with reduced holonomy which is not the case for $\Sigma$. But one could also charge the spinors under other global symmetries, introducing a twist of the spinor bundler which reduces the structure group and allows to find non-zero covariantly constant sections. 

In details, when a background connection $A$ for a global symmetry, take the $R$-symmetry, is turned on, the supercharges become sections of the product bundle of the spinor bundle with the vector bundle associated to the connection. Holonomies with respecto to $A$ ans $\omega$ are generically non-trivial but there might be some common subgroup of both the spin and $R$-symmetry group such that there are some sections that transform in opposite way under it. If this is the case, the structure group of the product bundle can be restricted to that subroup and the bundle decomposes into the direct sum of vector bundles associated to smaller irreducible representations of that subgroup. If the trivial rep appears in the decomposition, the associated sub-bundle admits non-zero flat sections, i.e charged Killing spinors. 

Take the example of the Riemann surface $\Sigma_g$. The structure group of the spinor bundle is $\rm{Spin}(2)\simeq \U(1)$, hence is a complex line bundle $\cL_\pm^\rm{Spin}$ with abelian connection $\omega$ associated to the $1$-dimensional rep of charge $\pm 1$. If we assume the existence of an $R$-symmetry $\U(1)_R$ such that supercharges have also charge $\pm 1$ under it, tgen tgere is a second line bundle $\cL^R_\pm$ such that 
\begin{equation}
  Q\in \Gamma(\cL_+^\rm{Spin}\oplus \cL_+^R)\qquad \overline{Q}\in \Gamma(\cL_-^\rm{Spin}\oplus \cL_-^R)
\end{equation}
The tensor product bundles have structure group $\rm{Spin}(2)\times \U(1)_R$ and connection $\omega+A$. If we restrict to the subgroup generated by elements of the form $(e^{i\theta},e^{-i\theta})$ and we choose $A=-\omega$, then the product bundle becomes topologically trivial and flat. All holonomies identically vanish and covariantly constant spinors can be found. 

More generally one can turn on fluxes for any global symmetry. In this case preserving susy also requires to set to zero the associated gaugino variations. 

\backmatter
\phantomsection
\addcontentsline{toc}{chapter}{References}
\bibliography{reps}

\end{document}
