\chapter{Rep theory of categorical symmetries}
How do categorical stymmetries act on and organiser thespectrum of operators in QFT? We are interested in genuine operators in the QFT, non-topological. The known idea is from representation theory as it usually applyies on invertible symmetries. Local op transform in some rep of a local symmetry, but also the same for lines which transform under rep of a one-form symmetry. This can be pushed to higher representations when zero form symmetries act on line defects, $2$-representations. 

This gives us an organising principle. But how do we do that for non-invertible symmetries? We have some (higher) fusion category which encompasses the action on some operators. Studying twisted sectors is key: consider local op attached to topological lines. We might have local operators attached to top lines, or line attached to topological surface etc. 

The answer, as alwyas, is conclusive for fusion categories $\cC$ in $d=2$: two answers
\begin{itemize}
	\item internal: representations of a tube algebra $\cA$. This is just the representation theory of an algebraic structure associated to the category.
	\item sandwitch construction: topolofical lines in $TV_\cC$. Via this construction all the twisted sector operators are attacched to a given topological line in the $3d$ theory.
\end{itemize}
Representations of the tube algebra are 
\begin{equation}
	\mathrm{Rep}(\cA)\equiv Z(\cC)
\end{equation}
where the right is the Drinfeld center of the category. 

Can this be extended to higher dimensions? Or also consider twisted secotr of nigher extended operators? Representtionf of $S^{d-n}$ Tube $n$-algebra, or $n$-dimensional operators in $TV_\cC$.
\begin{equation}
	\mathrm{nRep}\cA_{S^{d-n}}(\cA)\equiv \int_{S^{d-n}}\cC
\end{equation}

\section{Recap of $2d$}
The setup is as follows 
\begin{itemize}
	\item Spherical fusion category $\cC$: objects that correspond to topological lines $x,y,\ldots$. There is a notion of the dual with orientation reversal and object have fusion rules 
	\item $\pi_0(\cC)=$ Simple objects / isomorphisms. Fusion ring of category
	\item $x$-twistet local operator $x\in \pi_0(\cC)$
	\item $S^1$ tube action: lasso diagrams
\end{itemize}

The tube algebra is 
\begin{itemize}
	\item Finite semi-simple algevra $\cA_{S^1}(\cC)$
	\item Generated by morphisms in $\cC$ 
	\begin{equation}
		\psi:\cA\otimes x\rightarrow y\otimes \cA
	\end{equation}
	\item Product from horizontal composition
	\item Twisted local operators transform in irreps of this algebra. There is more structure
	\begin{equation}
		\mathrm{Rep}\cA_{S^1}(\cC)\quad \text{is a braided fusion category}
	\end{equation}
\end{itemize}
This is not an internal algebra of the category. 